### 2020-01-08 mstemc/index.tex
add the problem 1.1 and problem 1.2

% Intended LaTeX compiler: lualatex
\documentclass[koma,a4paper,utopia,12pt,listings-color,microtype,paralist,colorlinks,urlcolor=red]{org-article}
\usepackage{tikz}
\usetikzlibrary{arrows,decorations.pathmorphing,,backgrounds,positioning,fit,petri,calc,intersections,through,shapes.misc,mindmap,calendar,shadows,mindmap,calendar,graphdrawing,trees}
               \usepackage{tikz}
\author{Eason}
\date{\today}
\title{signals and systems chapter 1 problems}
\hypersetup{
 pdfauthor={Eason},
 pdftitle={signals and systems chapter 1 problems},
 pdfkeywords={},
 pdfsubject={summary of this post},
 pdfcreator={Emacs 26.3 (Org mode 9.2.6)},
 pdflang={English}}
\begin{document}

 \thispagestyle{empty}
 \begingroup
\begin{center}
 \vspace{\baselineskip}
 \textbf{\Huge signal and system chapter 1 problems} \par
 \vspace{2\baselineskip}  \newline
 \textbf{\large Eason Zhang with www.makesteamclear.com \par}
 \vspace{\baselineskip} \newline
  \today \par
 \vspace{\baselineskip}
  \vfill
 \setlength{\unitlength}{3pt}
 \includegraphics[width=0.5\textwidth]{/Users/chaolongzhang/Dropbox/mstemc_hugo/static/img/tikz/eulerEquation.pdf}
 \vfill \vspace{\baselineskip}
 \href{WWW.MAKESTEAMCLEAR.COM}{\Large WWW.MAKESTEAMCLEAR.COM} \par\newline
  makesteamclear is a free project, run by Eason Zhang, to make videos about STEAM in a more approachable way. If you find the contents in this article or the site or the youtube channel helpful, please consider \href{www.makesteamclear.com}{\ensuremath \heartsuit support me\ensuremath\heartsuit }, thanks \par
 \end{center}
 \endgroup

 \newpage \thispagestyle{empty} \textbf{copyright page} \newpage \tableofcontents \newpage
\hspace{0pt}\\

summary of this post
\section{Basic Problems With Answers}
\label{sec:org8f261cc}

\begin{prob}[]{prob:1.1}
Express each of the following complex numbers in \emph{Cartesian} form \(x+iy\):
\begin{equation*}
\begin{array}
\quad \frac{1}{2}e^{i\pi}  & \frac{1}{2}e^{-i\pi}  & e^{i\pi/2}  \\
 e^{-i\pi/2}         & e^{i5\pi/2}  & \sqrt{2}e^{i\pi/4}  \\
 \sqrt{2}e^{i9\pi/4} & \sqrt{2}e^{-9i\pi/4}  & \sqrt{2} e^{-i\pi/4}
\end{array}
\end{equation*}
\label{prob:1.1}
\end{prob}

\begin{prob}[]{prob:1.2}
Express each of the following complex numbers in \emph{polar} form (\(re^{i\theta},
-\pi < \theta \leq -\pi\)): \(5\),\(-2\),\(-3i\),\(\frac{1}{2} -
i\frac{\sqrt{3}}{2}\), \(1 + i\), \((1-i)^{2}\), \(i(1-i)\), \((1+i)/(1-i)\), \((\sqrt{2} +i\sqrt{2})/ (1 +i\sqrt{3})\)
\label{prob:1.2}
\end{prob}

Every complex number \(a+ib\) can be visualized in the complex plane
\(\mathbb{C}\). It can be viewed as the point with the coordinate \((a,b)\) in
the plane or as a vector starting from \(0,0\) to the point \((a,b)\).

Also every complex number \(a+ib\) can be represented in the exponential form
conveniently through the Euler's formula \(e^{i\alpha} = \cos\alpha +
i\sin\alpha\).

\begin{center}
\includegraphics[width=0.8\textwidth]{/Users/chaolongzhang/Dropbox/mstemc_hugo/static/img/tikz/eulerEquation.pdf}
\end{center}
\end{document}