### 2020-02-18 mstemc/mstemc.tex

% Intended LaTeX compiler: lualatex
\documentclass[koma,a4paper,utopia,12pt,listings-color,microtype,paralist,colorlinks,urlcolor=red]{org-article}
\usepackage{tikz}
\usetikzlibrary{arrows,decorations.pathmorphing,,backgrounds,positioning,fit,petri,calc,intersections,through,shapes.misc,mindmap,calendar,shadows,mindmap,calendar,graphdrawing,trees,shapes.misc,quotes,angles}
\definecolor{mycolor}{RGB}{139,0,0}
\usetheme{CambridgeUS}
\author{Eason}
\date{\today}
\title{A title}
\definecolor{links}{HTML}{8c1515}
\hypersetup{colorlinks=,linkcolor=,urlcolor=links}
\setbeamertemplate{itemize items}[default]
\setbeamertemplate{enumerate items}[default]
\setbeamertemplate{items}[default]
\setbeamercolor*{local structure}{fg=darkred}
\setbeamercolor{section in toc}{fg=darkred}
\setlength{\parskip}{\smallskipamount}
\institute[INST]{Institute\\\url{http://www.institute.edu}}
\titlegraphic{\includegraphics[height=1.5cm]{~/Dropbox/mstemc_hugo/static/img/icon-32.png}}
\hypersetup{
 pdfauthor={Eason},
 pdftitle={A title},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.3 (Org mode 9.2.6)},
 pdflang={English}}
\begin{document}

\framesubtitle{A subtitle}

\begin{itemize}
\item Bullet 1
\item Bullet 2
\item Bullet 3
\end{itemize}
\end{document}